---
title: "Vital signs"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
---

```{r setup, echo=FALSE, warning=FALSE, message = FALSE}
library(flexdashboard)
library(knitr)
library(rmarkdown)
library(dplyr)
library(lubridate)
library(qcc)
library(tseries)
library(forecast)
library(ggplot2)

opts_chunk$set(dev = "png", fig.path = "/home/greig/R-projects/Vital_Signs")

# load the data
ALOS.raw <- read.csv("/home/greig/R-projects/Vital_Signs/VS_ALOS.csv", header = TRUE, stringsAsFactors = FALSE, check.names = TRUE, strip.white = TRUE, na.strings = c("", "NULL", "NA"))

ALOS <- ALOS.raw

# declare the variables
ALOS$DDATE <- as.Date(ALOS$DDATE)
ALOS$EPISODE <- as.factor(ALOS$EPISODE)
ALOS$FACILITY <- as.factor(ALOS$FACILITY)
ALOS$ADMAGE <- as.integer(ALOS$ADMAGE)
ALOS$GENDER <- as.factor(ALOS$GENDER)
ALOS$ETHGROUP <-  as.factor(ALOS$ETHGROUP)
ALOS$DOMTLA <- as.factor(ALOS$DOMTLA)
ALOS$DOMDHB <- as.factor(ALOS$DOMDHB)
ALOS$DOMCODE <- as.factor(ALOS$DOMCODE)
ALOS$SPECIALTY <- as.factor(ALOS$SPECIALTY)
ALOS$ARRANGED <- as.factor(ALOS$ARRANGED)
ALOS$LOS <- as.integer(ALOS$LOS)
ALOS$LINE <- as.factor(ALOS$LINE)
ALOS$DRGCODE <- as.factor(ALOS$DRGCODE)
ALOS$PRINCIPLDIAGNOSISCODE <- as.factor(ALOS$PRINCIPLDIAGNOSISCODE)

ALOS <- ALOS %>% 
  mutate(Week = week(DDATE), Month = month(DDATE), Year = year(DDATE))

ALOS <- ALOS %>% 
  filter(DDATE < "2016-06-02 00:00:00")
```
#Introduction {.sidebar}
Vital signs (Part B) is focused on the analysis of patient flow. It provides detail and context regarding acute admissions. Some analysis of services that determine ALOS is also provided.

#ED

##Layout {.tabset .tabset-fade}

###Presentations
```{r ED-present, warning=FALSE, message = FALSE, comment="", results="hide"}

ED_arr <- read.csv("/home/greig/R-projects/Vital_Signs/ED_arr.csv", header = TRUE, stringsAsFactors = FALSE, check.names = TRUE, strip.white = TRUE, na.strings = c("", "NULL", "NA"))

ED_arr$EDArrTime <- as.Date(ED_arr$EDArrTime)
ED_arr <- ED_arr %>% 
  mutate(Year = year(EDArrTime), Month = month(EDArrTime), Week = week(EDArrTime))

ED_arr <- ED_arr %>% 
  filter(EDArrTime < "2016-06-02 00:00:00")

ED_count <- ED_arr %>% 
  group_by(Year, Week) %>% 
  summarise(Count = n()) %>% 
  ungroup()

c_ED <- select(ED_count, Count)
ewma(c_ED,  title = "Presentations / week for ED", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")
tsED <- ts(c_ED, frequency = 52, start = 2010)
tsED.dec1 <- decompose(tsED, type = "additive")
plot(tsED.dec1)

```

###Admissions
```{r ED-Admit, warning=FALSE, message = FALSE, comment="", results="hide"}

EM <- ALOS %>% 
  filter(ARRANGED == "Acute" & LINE == "Emergency")

totEM <- EM %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

c_totEM <- select(totEM, Count)
ewma(c_totEM,  title = "Admissions / week for Emergency Department", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")

tsEM <- ts(c_totEM, frequency = 52, start = 2010)
tsEM.dec1 <- decompose(tsEM, type = "additive")
plot(tsEM.dec1)
```

###ALOS 
```{r ED-ALOS, warning=FALSE, message = FALSE, comment="", results="hide"}

m_totEM <- select(totEM, Mean)
ewma(m_totEM, title = "ALOS for Emergency Department admissions", xlab = "Weeks since 1.Jan.2010", ylab = "ALOS")
```

###Forecast-6 months
```{r ED-forecast, warning=FALSE, message = FALSE, comment="", results="hide"}

EM2 <- ALOS %>% 
  filter(ARRANGED == "Acute" & LINE == "Emergency" & DDATE >= "2014-01-01 00:00:00")

totEM2 <- EM2 %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()


c_totEM2 <- select(totEM2, Count)

tsEM2 <- ts(c_totEM2, frequency = 52, start = 2014)

tsEM2.hwl <- HoltWinters(tsEM2, 
                         beta = NULL,
                         gamma = NULL, 
                         seasonal = "multiplicative")
tsEM2.fc3 <- forecast.HoltWinters(tsEM2.hwl,
                                  h=26)
plot(tsEM2.fc3)

```

###Forecast-1 yr
```{r ED-forecast2, warning=FALSE, message = FALSE, comment="", results="hide"}

tsEM2.fc4 <- forecast.HoltWinters(tsEM2.hwl,
                                  h=52)
plot(tsEM2.fc4)

```

#CSB

##Layout {.tabset .tabset-fade}

###ALOS 
```{r MCH-ALOS, warning=FALSE, message = FALSE, comment="", results="hide"}
MCH <- ALOS %>% 
  filter(LINE != "Emergency" & LINE != "Mental Health")

totMCH <- MCH %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

m_MCH <- select(totMCH, Mean)
ewma(m_MCH, title = "ALOS for CSB", xlab = "Weeks since 1.Jan.2010", ylab = "ALOS")

```

###Admissions
```{r MCH-Admit, echo=FALSE, warning=FALSE, message = FALSE, comment="", results="hide"}
c_MCH <- select(totMCH, Count)
ewma(c_MCH,  title = "Admissions / week for CSB", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")
tsMCH <- ts(c_MCH, frequency = 52, start = 2010)
tsMCH.dec1 <- decompose(tsMCH, type = "additive")
plot(tsMCH.dec1)
```

###Forecast-6 months
```{r MCH-FC, echo=FALSE, warning=FALSE, message = FALSE, comment="", results="hide"}
tsMCH.hwl <- HoltWinters(tsMCH, 
                         beta = NULL,
                         gamma = NULL, 
                         seasonal = "multiplicative")
tsMCH.fc3 <- forecast.HoltWinters(tsMCH.hwl,
                                  h=26)
plot(tsMCH.fc3)
```

###Forecast-1 yr
```{r MCH-FC2, echo=FALSE, warning=FALSE, message = FALSE, comment="", results="hide"}

tsMCH.fc4 <- forecast.HoltWinters(tsMCH.hwl,
                                  h=52)
plot(tsMCH.fc4)
```

#GenMed

##layout {.tabset .tabset-fade}

###ALOS
```{r GenMed_ALOS, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}
GenMed <- ALOS %>% 
  filter(ARRANGED == "Acute" & SPECIALTY == "General Internal Medical Services")

totGenMed <- GenMed %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

m_totGenMed <- select(totGenMed, Mean)
ewma(m_totGenMed, title = "ALOS for acute General Medicine", xlab = "Weeks since 1.Jan.2010", ylab = "ALOS")
```

###MAPU admissions
```{r GenMed_Admit, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

c_totGenMed <- select(totGenMed, Count)
ewma(c_totGenMed,  title = "MAPU admissions/week", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions/week")

tsGenMed <- ts(c_totGenMed, frequency = 52, start = 2010)
tsGenMed.dec1 <- decompose(tsGenMed, type = "additive")
plot(tsGenMed.dec1)

```

###MAPU discharges
```{r mapu-discharge, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

MAPU_disc <- GenMed %>% 
  filter(LOS<2)

MAPU_count <- MAPU_disc %>% 
  group_by(Year, Week) %>% 
  summarise(Count = n()) %>% 
  ungroup()


MAPU_percent <- right_join(totGenMed, MAPU_count, by=c("Year", "Week"))

MAPU_percent <- MAPU_percent %>% 
  mutate(Percent = round((Count.y/Count.x)*100, digits = 2))

ewma(MAPU_percent$Percent,  title = "MAPU % discharges/week", xlab = "Weeks since 1.Jan.2010", ylab = "Percentage discharges/week")

```

###Long Stay
```{r GenMed_LS, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

LS_GenMed <- GenMed %>% 
  filter(LOS>10)

LS_Med <- LS_GenMed %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

cLS_Med <- select(LS_Med, Count)
ewma(cLS_Med, title = "Number of patients admitted/week with LOS >10 days", xlab = "weeks since 1.Jan.2010", ylab = "admissions/week")
ts_cLS_Med <- ts(cLS_Med, frequency = 52, start = 2010)
tsLS_Med.dec1 <- decompose(ts_cLS_Med, type = "additive")
plot(tsLS_Med.dec1)

```

###Intermediate Stay
```{r GenMed_IS, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

IS_GenMed <- GenMed %>% 
  filter(LOS <= 10 & LOS >= 6)

IS_Med <- IS_GenMed %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

cIS_Med <- select(IS_Med, Count)
ewma(cIS_Med, title = "Number of patients admitted/week with LOS 6-9 days", xlab = "weeks since 1.Jan.2010", ylab = "admissions/week")
ts_cIS_Med <- ts(cIS_Med, frequency = 52, start = 2010)
tsIS_Med.dec1 <- decompose(ts_cIS_Med, type = "additive")
plot(tsIS_Med.dec1)

```

###Medium Stay
```{r GenMed_MS, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

MS_GenMed <- GenMed %>% 
  filter(LOS <= 5 & LOS >= 2)

MS_Med <- MS_GenMed %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

cMS_Med <- select(MS_Med, Count)
ewma(cMS_Med, title = "Number of patients admitted/week with LOS 2-5 days", xlab = "weeks since 1.Jan.2010", ylab = "admissions/week")
ts_cMS_Med <- ts(cMS_Med, frequency = 52, start = 2010)
tsMS_Med.dec1 <- decompose(ts_cMS_Med, type = "additive")
plot(tsMS_Med.dec1)

```

###1 Day Stay
```{r GenMed_vSS, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

vSS_GenMed <- GenMed %>% 
  filter(LOS>=0 & LOS <2)

vSS_GenMed <- vSS_GenMed %>% 
  filter(DDATE >= "2014-01-01 00:00:00")

vSS_Med <- vSS_GenMed %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

vcSS_Med <- select(vSS_Med, Count)
ewma(vcSS_Med, title = "Number of patients admitted/week with LOS <=24hr days", xlab = "weeks since 1.Jan.2014", ylab = "admissions/week")
vts_cSS_Med <- ts(vcSS_Med, frequency = 52, start = 2014)
vtsSS_Med.dec1 <- decompose(vts_cSS_Med, type = "additive")
plot(vtsSS_Med.dec1)

```

#GenSurg

##layout{.tabset .tabset-fade}

###ALOS - service
```{r GenSurg-ALOS, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}
GenSurg <- ALOS %>% 
  filter(SPECIALTY == "General Surgical Services")

totGenSurg <- GenSurg %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

m_totGenSurg <- select(totGenSurg, Mean)
ewma(m_totGenSurg, title = "ALOS for acute General Surgery", xlab = "Weeks since 1.Jan.2010", ylab = "ALOS")
c_totGenSurg <- select(totGenSurg, Count)
```

###Total Admissions
```{r GenSurg-Admit, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

c_totGenSurg <- select(totGenSurg, Count)
ewma(c_totGenSurg,  title = "Admissions / week for acute General Surgery", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")

tsGenSurg <- ts(c_totGenSurg, frequency = 52, start = 2010)
tsGenSurg.dec1 <- decompose(tsGenSurg, type = "additive")
plot(tsGenSurg.dec1)

```

###Acute Admissions
```{r GenSurg-Acute_Admit, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

SGenSurg <- ALOS %>% 
  filter(ARRANGED == "Acute" & SPECIALTY == "General Surgical Services")

SSGenSurg <- SGenSurg %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

c_SSGenSurg <- select(SSGenSurg, Count)
ewma(c_SSGenSurg,  title = "Acute Gen Surgery admissions / week", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")

ts_SSGenSurg <- ts(c_SSGenSurg, frequency = 52, start = 2010)
ts_SSGenSurg.dec1 <- decompose(ts_SSGenSurg, type = "additive")
plot(ts_SSGenSurg.dec1)

```

#Ortho

##layout {.tabset .tabset-fade}

###ALOS - service
```{r Ortho-ALOS, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

Ortho <- ALOS %>% 
  filter(SPECIALTY == "Orthopaedic Services")

totOrtho <- Ortho %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

m_totOrtho <- select(totOrtho, Mean)
ewma(m_totOrtho, title = "ALOS for Orthopaedic Surgery", xlab = "Weeks since 1.Jan.2010", ylab = "ALOS")

```

###Total Admissions
```{r Ortho-Admit, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

c_totOrtho <- select(totOrtho, Count)
ewma(c_totOrtho,  title = "Admissions / week for Orthopaedic Surgery", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")

tsOrtho <- ts(c_totOrtho, frequency = 52, start = 2010)
tsOrtho.dec1 <- decompose(tsOrtho, type = "additive")
plot(tsOrtho.dec1)

```

###Acute Admissions
```{r Ortho-Acute, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

acuteOrtho <- ALOS %>% 
  filter(ARRANGED == "Acute" & SPECIALTY == "Orthopaedic Services")

tot_acuteOrtho <- acuteOrtho %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

c_totacuteOrtho <- select(tot_acuteOrtho, Count)
ewma(c_totacuteOrtho,  title = "Admissions / week for acute Orthopaedic Surgery", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")

tsacuteOrtho <- ts(c_totacuteOrtho, frequency = 52, start = 2010)
tsacuteOrtho.dec1 <- decompose(tsacuteOrtho, type = "additive")
plot(tsacuteOrtho.dec1)

```


###Elective Admissions
```{r Ortho-Elective, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

electOrtho <- ALOS %>% 
  filter(ARRANGED != "Acute" & SPECIALTY == "Orthopaedic Services")

tot_electOrtho <- electOrtho %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

c_totelectOrtho <- select(tot_electOrtho, Count)
ewma(c_totelectOrtho,  title = "Admissions / week for elective Orthopaedic Surgery", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")

tselectOrtho <- ts(c_totelectOrtho, frequency = 52, start = 2010)
tselectOrtho.dec1 <- decompose(tselectOrtho, type = "additive")
plot(tselectOrtho.dec1)

```

#ElderHealth

##layout{.tabset .tabset-fade}

###ALOS-service
```{r EH-ALOS, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

EH <- ALOS %>% 
  filter(LINE == "Elderly")

totEH <- EH %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

m_totEH <- select(totEH, Mean)
ewma(m_totEH, title = "ALOS for Elder Health", xlab = "Weeks since 1.Jan.2010", ylab = "ALOS")

tsEHm <- ts(m_totEH, frequency = 52, start = 2010)
tsEHm.dec1 <- decompose(tsEHm, type = "additive")
plot(tsEHm.dec1)

```

###Admissions-service
```{r EH-Admit, echo=FALSE, warning=FALSE, message = FALSE, results="hide"}

c_totEH <- select(totEH, Count)
ewma(c_totEH,  title = "Admissions / week for Elder Health", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")

tsEH <- ts(c_totEH, frequency = 52, start = 2010)
tsEH.dec1 <- decompose(tsEH, type = "additive")
plot(tsEH.dec1)

```

#Mental Health

##layout{.tabset .tabset-fade}

###ALOS-service
```{r MH-ALOS, echo=FALSE, warning=FALSE, message = FALSE, comment="", results='hide'}
MH <- ALOS %>% 
  filter(LINE == "Mental Health")

totMH <- MH %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

m_totMH <- select(totMH, Mean)
ewma(m_totMH, title = "ALOS for Mental Health", xlab = "Weeks since 1.Jan.2010", ylab = "ALOS")
c_totMH <- select(totMH, Count)

```

###Admissions-service
```{r MH-Admit, echo=FALSE, warning=FALSE, message = FALSE, comment="", results='hide'}

c_totMH <- select(totMH, Count)
ewma(c_totMH,  title = "Admissions / week for Mental Health", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")

tsMH <- ts(c_totMH, frequency = 52, start = 2010)
tsMH.dec1 <- decompose(tsMH, type = "additive")
plot(tsMH.dec1)

```

#Child Health

##layout{.tabset .tabet-fade}

###ALOS-acute
```{r paeds-alos, echo=FALSE, warning=FALSE, message = FALSE, comment="", results='hide'}

Paed <- ALOS %>% 
  filter(ARRANGED == "Acute" & LINE == "Paediatric")

totPaed <- Paed %>% 
  group_by(Year, Week) %>% 
  summarise(Mean = mean(LOS, na.rm = TRUE), Count = n()) %>% 
  ungroup()

m_totPaed <- select(totPaed, Mean)
ewma(m_totPaed, title = "ALOS for acute Child Health admissions", xlab = "Weeks since 1.Jan.2010", ylab = "ALOS")
```

###Admissions-acute
```{r paeds-admit, echo=FALSE, warning=FALSE, message = FALSE, comment="", results='hide'}
c_totPaed <- select(totPaed, Count)
ewma(c_totPaed,  title = "Admissions / week for acute Child Health admissions", xlab = "Weeks since 1.Jan.2010", ylab = "Admissions / week")

tsPaed <- ts(c_totPaed, frequency = 52, start = 2010)
tsPaed.dec1 <- decompose(tsPaed, type = "additive")
plot(tsPaed.dec1)

```